StreamFind for mass spectrometry

### 2024-04-29 dev/dev_ideas/ms_overview.Rmd
Major update of MassSpecEngine Semantic

Within the MassSpecData class object, standardized methods for plotting data are also available. All plot methods use data that can be obtained with get methods, meaning that the user is able to re-plot the data with a different approach/style. For instance, the `get_spectra_tic()` method can be used to obtain the total ion chromatogram (TIC) as a `data.table` with columns *analysis*, *rt* and *intensity*, as shown below for the first analysis. The same TIC data can be directly plotted using the `plot_tic()` method, as shown below.

@@

```{r plot-tic} 
ms$plot_tic(analyses = 1)
```

@@

```{r plot-bpc-static}
ms$plot_bpc(interactive = FALSE)
```

@@

Another plotting option for raw spectra is the `plot_xic()` method as shown below. This method returns a two row plot with traces intensity vs retention time on top and *m/z* vs retention time below. A `targetsMark` can be defined to illustrate the area of expected *m/z* and retention time targets. Below, the targets1 rt and mz columns with narrower ppm and sec deviations (i.e., 5 ppm and 10 seconds, respectively) is used to show the area where Carbamazepin-d10 is expected for potential identification.

```{r plot-xic}
ms$plot_xic(analyses = 1:2, mz = targets1[2,],
  targetsMark = targets1[2, c("rt", "mz")], ppmMark = 5, secMark = 10)
```

@@

The extracted ion chromatogram (EIC) can be obtained with `get_spectra_eic()` and plotted with `plot_eic()` method. The interface is very similar as the previous methods, with analyses and make targets arguments used to specify each analyses to parse the data and which traces to retrieve.

@@

```{r plot-eic-01}
ms$plot_eic(mz = targets1, interactive = FALSE)
```

```{r plot-eic-02}
ms$plot_eic(mz = targets1, colorBy = "analyses")
```

@@

```{r plot-ms1}
# In the plot, the Carbamazepin-d10 and Diuron-d6 isotopes can be distinguished
ms$plot_ms1(analyses = 3,
  mz = data.frame("mzmin" = c(247, 239), "mzmax" = c(257, 245)),
  rt = data.frame("rtmin" = targets1$rt - 1, "rtmax" = targets1$rt + 1),
  mzClust = 0.005, interactive = FALSE
)
```

@@

```{r plot-ms2-01}
# MS2 data plot of Diuron-d6 in all analyses
ms$plot_ms2(mz = targets1[2 ,], 
  mzClust = 0.005, minIntensity = 50,
  colorBy = "analyses", verbose = FALSE)
```

```{r plot-ms2-02}
# MS2 data plot of both targets in analyses 3
# Note that "minIntensity" argument was increased to 500 counts, resulting in a cleaner averaged spectrum
ms$plot_ms2(analyses = 3, mz = targets1, 
  mzClust = 0.005, minIntensity = 500,
  colorBy = "targets", verbose = FALSE)
```

>>>>

Within the MassSpecData class object, standardized methods for plotting data are also available. All plot methods use data that can be obtained with get methods, meaning that the user is able to re-plot the data with a different approach/style. For instance, the `get_spectra_tic()` method can be used to obtain the total ion chromatogram (TIC) as a `data.table` with columns *analysis*, *rt* and *intensity*, as shown below for the first analysis. The same TIC data can be directly plotted using the `plot_spectra_tic()` method, as shown below.

@@

```{r plot-tic} 
ms$plot_spectra_tic(analyses = 1)
```

@@

```{r plot-bpc-static}
ms$plot_spectra_bpc(interactive = FALSE)
```

@@

Another plotting option for raw spectra is the `plot_spectra_xic()` method as shown below. This method returns a two row plot with traces intensity vs retention time on top and *m/z* vs retention time below. A `targetsMark` can be defined to illustrate the area of expected *m/z* and retention time targets. Below, the targets1 rt and mz columns with narrower ppm and sec deviations (i.e., 5 ppm and 10 seconds, respectively) is used to show the area where Carbamazepin-d10 is expected for potential identification.

```{r plot-xic}
ms$plot_spectra_xic(analyses = 1:2, mz = targets1[2,],
  targetsMark = targets1[2, c("rt", "mz")], ppmMark = 5, secMark = 10)
```

@@

The extracted ion chromatogram (EIC) can be obtained with `get_spectra_eic()` and plotted with `plot_spectra_eic()` method. The interface is very similar as the previous methods, with analyses and make targets arguments used to specify each analyses to parse the data and which traces to retrieve.

@@

```{r plot-eic-01}
ms$plot_spectra_eic(mz = targets1, interactive = FALSE)
```

```{r plot-eic-02}
ms$plot_spectra_eic(mz = targets1, colorBy = "analyses")
```

@@

```{r plot-ms1}
# In the plot, the Carbamazepin-d10 and Diuron-d6 isotopes can be distinguished
ms$plot_spectra_ms1(analyses = 3,
  mz = data.frame("mzmin" = c(247, 239), "mzmax" = c(257, 245)),
  rt = data.frame("rtmin" = targets1$rt - 1, "rtmax" = targets1$rt + 1),
  mzClust = 0.005, interactive = FALSE
)
```

@@

```{r plot-ms2-01}
# MS2 data plot of Diuron-d6 in all analyses
ms$plot_spectra_ms2(mz = targets1[2 ,], 
  mzClust = 0.005, minIntensity = 50,
  colorBy = "analyses", verbose = FALSE)
```

```{r plot-ms2-02}
# MS2 data plot of both targets in analyses 3
# Note that "minIntensity" argument was increased to 500 counts, resulting in a cleaner averaged spectrum
ms$plot_spectra_ms2(analyses = 3, mz = targets1, 
  mzClust = 0.005, minIntensity = 500,
  colorBy = "targets", verbose = FALSE)
```
